---
title: "ReactomePA Alison"
author: "Joana Bittencourt Silvestre"
date: "24/11/2020"
output: html_document
---

First, let's load the package
```{r}
library(ReactomePA)
library(dplyr)
library(tidyr)
library(stringi)
library(stringr)
library(clusterProfiler)
library(lubridate)
```

Set names of directories, input and output files
```{r}
wd="/Volumes/SS-Bioinf/Projects/Vignir/Joana_Amy/05_Analysis/noNeut/Nov2/"
```

```{r}
#read all files from chosen directory
#chose the ones that have a certain pattern on them (usually "DESeq2")
#take always the filtered ones
getFilesList<-function(wd){
  setwd(wd)
  listFiles<-Sys.glob("*DESeq2*.csv")
  filtMask<-!str_detect(listFiles, "filt")
  notEmpty<-file.size(listFiles) > 0
  listFiles<-listFiles[filtMask & notEmpty] 
  return(listFiles)
}
```


```{r}
getTitles<-function(lFiles){
  lTitles<-list()
  for (i in 1:length(lFiles)){
    f<-lFiles[[i]]
    t<-strsplit(f,"\\.")[[1]]
    t1 <- gsub("_", " ", t[[2]])
    lTitles[[i]] <- t1
    i <- i + 1
  }
  return(lTitles)
}
```

```{r}
getGeneList<-function(listFiles, threshold){
  geneList<-list()
  i=1
  for(file in listFiles){
    df<-read.csv(paste0(wd,file))
    if(threshold<0){
      df = df %>% dplyr::filter(log2FoldChange < threshold & padj <0.05 ) %>%
      dplyr::select(entrez_id, log2FoldChange) %>%
      drop_na() %>%
      arrange(log2FoldChange)
        sort
    } else {
      df= df %>% dplyr::filter(log2FoldChange > threshold & padj <0.05 ) %>%
      dplyr::select(entrez_id, log2FoldChange) %>%
      drop_na()%>%
      arrange(log2FoldChange)
    }
    
    genes<-as.vector(df$log2FoldChange)
    names(genes)<-as.character(df$entrez_id)
    geneList[[i]]<-genes
    i=i+1
  }
  return(geneList)
}

```



Get all the lists you need
```{r}
filesList<-getFilesList(wd)
geneList_POS=getGeneList(filesList, 1.5)
geneList_NEG=getGeneList(filesList, -1.5)
titlesList_half<-getTitles(filesList)
```

```{r}
addWordToTitles=function(listTitles, word){
  newList=list()
  i=1
  for(title in titlesList_half){
    newTitle=paste(title, word)
    newList[[i]]=newTitle
    i=i+1
  }
  return(newList)
}
```

```{r}
titles_Pos=addWordToTitles(titlesList_half, "Pos")
titles_Neg=addWordToTitles(titlesList_half, "Neg")
```


now, combine gene list and title list
```{r}
geneList=c(geneList_POS, geneList_NEG)
titlesList=c(titles_Pos, titles_Neg)
```




Don't forget that you must download org.Mm.eg.db or org.Hs.eg.db yourself!
And also load the package!

Now do the enrichment and establish a p value cuttof
function takes names of genes instead of LFC, and this is important because the LFC can be used later (justa call geneList[[x]] for whatever you need LFC)
```{r}
library(org.Mm.eg.db)

#function to do the enrichment for several lists of genes
getEnPath<-function(genes){
  #use that numeric list that we have just obtained
  listRes=list()
  i=1
  for(genes in geneList){
    enPath<-enrichPathway(gene=names(geneList[[i]]), organism = "mouse", pvalueCutoff=0.05, readable=T)
    listRes[[i]]<-enPath
    i=i+1
  }
  return(listRes)
}

```


Now use the above created function to create list of enriched pathways
```{r}
enPathList<-getEnPath(geneList)
```


Let's find out which enPath is empty
```{r}
listEval<-lapply(enPathList, `[`, 1 ) %>% lapply(`[[`,1)
#And use that info to get that result out of enPathList, titles and gene list
enPathList.filt<-enPathList[!is.na(listEval)]
geneList.filt<-geneList[!is.na(listEval)]
titlesList.filt<-titlesList[!is.na(listEval)]
```

And now everything is kind of ready to test the graphs, so I'll do all results per graph type
you must use print to plot your graph in order for png to work in a loop
```{r}
i=1
for(e in enPathList.filt){
  fname<-paste0(wd,"reactomePA_barPlot_",titlesList.filt[[i]],today(),".png")
  png(fname, width = 35, height = 10, units = "cm", res=300)
  print(barplot(enPathList.filt[[i]], showCategory=8))
  dev.off( )
  i=i+1
}
```


```{r}
i=1
for(e in enPathList.filt){
  fname<-paste0(wd,"reactomePA_dotPlot_",titlesList.filt[[i]],today(),".png")
  png(fname, width = 25, height = 10, units = "cm", res=300)
  print(dotplot(enPathList.filt[[i]], showCategory=15))
  dev.off()
  i=i+1
}

```

```{r}
i=1
for(e in enPathList.filt){
  fname<-paste0(wd,"reactomePA_emapplot_",titlesList.filt[[i]],today(),".png")
  png(fname, width = 35, height = 25, units = "cm", res=300)
  print(emapplot(enPathList.filt[[i]]))
  dev.off()
  i=i+1
}
```

```{r}
i=1
for(e in enPathList.filt){
  fname<-paste0(wd,"reactomePA_cnetplot_",titlesList.filt[[i]],today(),".png")
  png(fname, width = 35, height = 25, units = "cm", res=300)
  print(cnetplot(enPathList.filt[[i]], categorySize="pvalue", foldChange=geneList.filt[[i]]))
  dev.off()
  i=i+1
}
```
Now using cluster profiles for some cooler graphs
Num needs to have NAME!!!!

```{r}
require(clusterProfiler)
names(geneList.filt)<-titlesList.filt
res <- compareCluster(lapply(geneList.filt, names), fun="enrichPathway")
fname<-paste0(wd,"reactomePA_clusterProf_posAndNeg_",today(),".svg")
dev.new()
svg(fname)
dotplot(res, font.size=10)
dev.off()
```



```{r}
sessionInfo()
```
